---
title: "Liberate Your Time: Accelerate with R - Wrangle Data"
author: "Thomas Sharpe,"
date: "2023-05-24"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, fig.width = 12, fig.height = 6)

knitr::opts_knit$set(root.dir = 'C:\\Users\\Thomas Sharpe\\Documents\\thomasIR\\air2023')

```

## Wrangling Data

As this presentation is not intended to be an R tutorial, this section will be brief. In short, the dplyr functionality is great. Readable and quick. Personally preferred over alternatives in Python (e.g. pandas). Just me though! :)

```{r, eval = FALSE}

library(dplyr)

#This framework is useful super often.
newDF <-
  df %>%
  filter() %>%
  group_by() %>%
  summarise() %>%
  arrange()

```

## RStudio Cope Snippets

Tools -> Edit Code Snippets

Within that menu:

snippet dplyrTemplate (call it whatever you want)
	df <-
		filter() %>%
		group_by() %>%
		summarise() %>%
		arrange()
		
Most IDE's have something like this. I have used it more for SQL than R though truthfully. Institution-specific common queries or common starting points for queries. <b>Huge time saver </b> ( beats AI, for now :) ) 


```{r, eval = FALSE}

# Start typing the name of snippet, it will show up in your suggestions after three characters entered, press tab and bam.

df <-
  filter() %>%
  group_by() %>%
  summarise() %>%
  arrange()

```


